---
title: "Predictive Model for Movie Revenue Estimation and Decision Support"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  #pdf_document:
    #toc: yes
    #toc_depth: '3'
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

Detailed Problem Statement:
The movie industry is a dynamic and complex environment where various factors influence a movie's success, particularly its revenue. This project focuses on developing a robust "Predictive Model for Movie Revenue Estimation and Decision Support." In an era of high uncertainty and financial risk in the film production business, understanding the determinants of movie revenue is paramount for informed decision-making, investment strategies, and marketing planning.

Movies are a unique combination of artistic expression and commercial ventures. To that end, this research investigates a multitude of features that contribute to the financial performance of a movie. These include aspects like genres, original languages, overview, popularity, production companies, release date, budget, runtime, status, tagline, vote average, vote count, credits, keywords, poster path, backdrop path, and recommendations. Each of these factors is scrutinized meticulously to unravel their individual and collective influence on movie revenue.

By conducting a comprehensive analysis of these features, this research aims to build a regression model (Linear, Decision Tree, Support Vector Machine) based on valuable insights through identifying patterns, relationships, and correlations. The findings of this study are expected to have substantial implications for movie production companies, investors, and stakeholders. They can leverage the predictive model developed in this project to make informed decisions regarding their movie investments, marketing strategies, and financial planning, ultimately contributing to the success and profitability of their ventures in the film industry.

# Understanding the Data
The movie dataset under analysis originates from The Movie Database (TMDb) API, replacing the original IMDb dataset due to a DMCA takedown request. This comprehensive dataset spans plot details, cast and crew information, budget, revenues, and more for thousands of films. The transition to TMDb includes improvements such as full credits for both cast and crew, listing actors in credit order, and updated revenue figures for greater accuracy. This particular dataset has been made
available on [Kaggle](<https://www.kaggle.com/datasets/akshaypawar7/millions-of-movies>), which was the
source of reference. The entire dataset contains movies from 1950s to 2020s. We have more than a million rows but for computational limitations, we will analyse using a sample of the data.
The dataset sparks curiosity about predicting movie success and categorizing films by type, utilizing crew job titles for insights. Additionally, it prompts exploration into the dynamics between major film studios and independents, potentially through clustering analysis.

Reference guide for the columns in the dataset:

1. title: Movie or show name
2. genres: Content categories or themes
3. original_language: Language of the original content
4. popularity: Measure of audience interest
5. release_date: Date of public availability
6. budget: Cost of production
7. revenue: Income generated from content
8. runtime: Duration of the content
9. status: Current release status
10. vote_average: Average audience ratings
11. vote_count: Number of audience votes
12.trailer_views: Number of trailer views
13. trailer_likes: Number of trailer likes

**Importing the Movies Data**

```{r}
movies = read.csv("Movie_data.csv")
#head(movies)
summary(movies)
movie = movies # Creating a copy of data frame
```

**Removing Unwanted Columns**

```{r}
library(dplyr)
remove_col <- c(1,5,7,13,16,17,18,19,20)
movie <- movie[, -remove_col]
head(movie)
```

**Summary of the data**

```{r}
new_column_names <- c("Title", "Genre", "Language", "Popularity", "Release_Date","Budget","Revenue","Runtime","Status","Avg_Vote","Vote_Count","Trailer_Views","Trailer_Likes")
colnames(movie) <- new_column_names
summary(movie)
```

Inferences:
1. Budget of Movies are having value 0, which is not possible (Remove rows with budget 0)
2. Revenue has negative values which is again not possible (Remove rows with revenue 0 or negative)
3. Runtime of Movies are having value 0, which is not possible (Remove rows with runtime 0)
4. Check for outliers and influential points in all the columns.

**Checking for missing values in the data**

```{r}
library(ggplot2)

# Calculate the percentage of missing values in each column
column_na_percentages <- colSums(is.na(movie)) / nrow(movie) * 100

# Set a smaller font size
smaller_font_size <- 0.8

# Create a bar plot for the percentage of missing values in each column
barplot(column_na_percentages, main = "Percentage of Missing Values in Each Column",
        xlab = "Columns", ylab = "Percentage of Missing Values",
        col = "lightcoral", border = "black", ylim = c(0, 2),
        names.arg = names(column_na_percentages), las = 2, cex.names = smaller_font_size)

print("Rows in Dataset before dropping rows with NA Values")
print(nrow(movie))

movie <- movie[complete.cases(movie), ]

print("Rows in Dataset After dropping rows with NA Values")
print(nrow(movie))
```

Removed rows with NA Values

***Data Cleaning***

```{r}
print("Rows in Dataset before cleaning")
print(nrow(movie))

# 1. Remove rows with Budget less than or equal to 0
movie <- movie[movie$Budget > 0, ]

# 2. Remove rows with Revenue less than or equal to 0
movie <- movie[movie$Revenue > 0, ]

# 3. Remove rows with Runtime less than or equal to 0
movie <- movie[movie$Runtime > 0, ]

print("Rows in Dataset After cleaning")
print(nrow(movie))
```

Inferences:

1. Budget of Movies are having value 0, which is not possible (Remove rows with budget 0)

2. Revenue has negative values which is again not possible (Remove rows with revenue 0 or negative)

3. Runtime of Movies are having value 0, which is not possible (Remove rows with runtime 0)

4. Check for outliers and influential points in all the columns.

5. Runtime had 0.004% missing values.

6. These values were dropped as number of missing value is small.

**Data Visulisation**

**Boxplots for Selected Columns**

```{r}
selected_columns <- c("Popularity", "Budget", "Revenue", "Runtime", "Avg_Vote", "Vote_Count", "Trailer_Views","Trailer_Likes")

box_plots_list <- list()

for (col in selected_columns) {
  box_plot <- ggplot(movie, aes(y = get(col))) +
    geom_boxplot(fill = "lightblue", color = "black", alpha = 0.7) +
    labs(title = paste("Box Plot for", col), y = col) +
    theme_minimal() +
    theme(plot.title = element_text(size = 15)) +
    theme(axis.title = element_text(size = 12)) +
    theme(axis.text = element_text(size = 10))
  box_plots_list[[col]] <- box_plot
}
for (col in selected_columns) {
  print(box_plots_list[[col]])
}
```

**Scatterplot for Revenue**

```{r}
predictors <- c("Trailer_Likes", "Trailer_Views", "Budget", "Avg_Vote","Vote_Count","Runtime")
scatter_plots <- lapply(predictors, function(var) {
  ggplot(movie, aes(x = get(var), y = Revenue, color = Revenue)) +
    geom_point() +
    labs(title = paste("Scatter plot of", var, "vs. Revenue"),
         x = var,
         y = "Revenue",
         color = "Revenue")
})
scatter_plots
```

Inferences:

Trailer Likes and Trailer views have linear relationship with the Output/Dependent variable "Revenue".

**Scatterplot for Popularity**

```{r}
predictors <- c("Trailer_Likes", "Trailer_Views", "Vote_Count", "Budget", "Runtime", "Avg_Vote")

scatter_plots<- lapply(predictors, function(var) {
  ggplot(movie, aes(x = get(var), y = Popularity, color = Popularity)) +
    geom_point() +
    labs(
      title = paste("Scatter plot of", var, "vs. Popularity"),
      x = var,  
      y = "Popularity",
      color = "Popularity"
    )
})

scatter_plots

```

**Scatterplot for Popularity**

```{r}
predictors <- c("Trailer_Likes", "Trailer_Views", "Vote_Count", "Budget", "Runtime", "Avg_Vote")

scatter_plots<- lapply(predictors, function(var) {
  ggplot(movie, aes(x = get(var), y = Popularity, color = Popularity)) +
    geom_point() +
    labs(
      title = paste("Scatter plot of", var, "vs. Popularity"),
      x = var,  
      y = "Popularity",
      color = "Popularity"
    )
})
scatter_plots

```

>>>>>>> Stashed changes
**Visualization of Correlation Matrix using Heatmap**

```{r}
numeric_cols <- sapply(movie, is.numeric)
cor_matrix <- cor(movie[, numeric_cols])
cor_matrix
library(corrplot)
corrplot(cor_matrix, method = "color")

```

Inferences:

1. Trailer Likes and Trailer views has high correlation with "Revenue".

2. Vote Count and Budget have moderate correlation with "Revenue".

3. Popularity, Runtime and Average Vote have poor correlation with "Revenue".

**Pairwise Scatterplots**

```{r}
library(GGally)
numeric_vars_subset <- c("Revenue", "Budget", "Popularity", "Runtime")

ggpairs(movie[, numeric_vars_subset])

```

**Correlation with Target variable and visualization of it**

```{r}
cor_with_revenue <- sapply(movie[, numeric_cols], function(x) cor(x, movie$Revenue))
cor_with_revenue
barplot(cor_with_revenue, names.arg = names(cor_with_revenue), col = "skyblue", main = "Correlation with Revenue")

```

**Stepwise Forward Feature Selection**
```{r}
library(MASS)

# Define your initial model
initial_model <- lm(Revenue ~ 1, data = movie)  # Simple model with intercept only

# Perform stepwise forward selection
final_model <- stepAIC(initial_model, direction = "forward", scope = list(lower = ~1, upper = ~Popularity + Budget + Runtime + Avg_Vote + Vote_Count + Trailer_Views + Trailer_Likes), data = movie)

# Display the final model summary
summary(final_model)

```

The features given in the final model using forward feature selection are Trailer_Likes, Trailer_Views, Vote_Count, Budget, Runtime ,Avg_Vote , popularity but we can see that popularity is not significant hence we will drop popularity and build the model again.

```{r}
lm_model_usingFFS <- lm(Revenue ~ Trailer_Likes + Trailer_Views + Vote_Count + Budget + Runtime + Avg_Vote, data = movie)

# Summarize the model
summary(lm_model_usingFFS)
```

In the above model we can see that all the input features are significant and we have a good Multiple R-Squared value. 
The p-value for the model is also significant indicating the model is a significant model.
The Residual Std Error is around 5.4 Million which seems a little high as we cannot exactly predict the Revenue of a Movie as it depends on various other factors.
```{r}
library(caret)

# Set the seed for reproducibility
set.seed(123)

# Define the predictors and response variable
predictors <- c("Trailer_Likes", "Trailer_Views", "Vote_Count", "Budget", "Runtime", "Avg_Vote")
response <- "Revenue"

# Create a training and testing dataset
train_index <- createDataPartition(movie$Revenue, p = 0.8, list = FALSE)
train_data <- movie[train_index, ]
test_data <- movie[-train_index, ]

# Fit the model on the training data
model <- lm(paste(response, "~", paste(predictors, collapse = " + ")), data = train_data)

# Make predictions on the test data
predictions <- predict(model, newdata = test_data)

# Calculate RMSE
rmse <- sqrt(mean((test_data$Revenue - predictions)^2))

# Print the RMSE
cat("Root Mean Squared Error (RMSE):", rmse, "\n")

```

```{r}
lm_model <- lm(Revenue ~ Trailer_Views+Trailer_Likes, data = movie)

# Summarize the model
summary(lm_model)
```

```{r}
library(caTools)

# Set a random seed for reproducibility
set.seed(123)

# Split the data into training (80%) and testing (20%) sets
split <- sample.split(movie$Revenue, SplitRatio = 0.8)
train_data <- subset(movie, split == TRUE)
test_data <- subset(movie, split == FALSE)

# Fit linear regression model on the training data
lm_model <- lm(Revenue ~ Trailer_Views+Trailer_Likes, data = train_data)

# Make predictions on the test data
predictions <- predict(lm_model, newdata = test_data)

# Calculate Root Mean Squared Error (RMSE)
rmse <- sqrt(mean((test_data$Revenue - predictions)^2))
cat("Root Mean Squared Error (RMSE):", rmse, "\n")
```

```{r}
mean_revenue <- mean(movie$Revenue, na.rm = TRUE)
median_revenue <- median(movie$Revenue, na.rm = TRUE)
quantiles_revenue <- quantile(movie$Revenue, c(0.25, 0.5, 0.75), na.rm = TRUE)

# Print the results
cat("Mean Revenue:", mean_revenue, "\n")
cat("Median Revenue:", median_revenue, "\n")
cat("Quantiles (25%, 50%, 75%):", quantiles_revenue, "\n")
```
