---
title: "Predictive Model for Movie Revenue Estimation and Decision Support"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  #pdf_document:
    #toc: yes
    #toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Importing the data set, include=F} 
movie = read.csv("movies.csv")
head(movie)
```


Detailed Problem Statement:
The movie industry is a dynamic and complex environment where various factors influence a movie's success, particularly its revenue. This project focuses on developing a robust "Predictive Model for Movie Revenue Estimation and Decision Support." In an era of high uncertainty and financial risk in the film production business, understanding the determinants of movie revenue is paramount for informed decision-making, investment strategies, and marketing planning.

Movies are a unique combination of artistic expression and commercial ventures. To that end, this research investigates a multitude of features that contribute to the financial performance of a movie. These include aspects like genres, original languages, overview, popularity, production companies, release date, budget, runtime, status, tagline, vote average, vote count, credits, keywords, poster path, backdrop path, and recommendations. Each of these factors is scrutinized meticulously to unravel their individual and collective influence on movie revenue.

By conducting a comprehensive analysis of these features, this research aims to build a regression model (Linear, Decision Tree, Support Vector Machine) based on valuable insights through identifying patterns, relationships, and correlations. The findings of this study are expected to have substantial implications for movie production companies, investors, and stakeholders. They can leverage the predictive model developed in this project to make informed decisions regarding their movie investments, marketing strategies, and financial planning, ultimately contributing to the success and profitability of their ventures in the film industry.

**Importing and Summary of the Stroke Data**

